Boutiques: a flexible framework for automated application integration in computing platforms

### 2017-11-02 boutiquespaper.tex
A few fixes, added paragraph on BIDS apps, added affiliations, removed Flywheel

\author{Tristan Glatard, Gregory Kiar, Tristan Aumentado-Armstrong,
  Natacha Beck, Pierre Bellec, Remi Bernard, Sorina Camarasu-Pop,
  Fr\'ed\'eric Cervenansky, Samir Das,  Rafael Ferreira da Silva,
  Guillaume Flandin, John Flavin, Pascal Girard, Krzysztof
  J. Gorgolewski, Charles R.G. Guttmann, Val\'erie Hayot-Sasson,
  Nathaniel Kofalt, Pierre-Olivier Quirion, Pierre Rioux,
  Marc-\'Etienne Rousseau, Gunnar Schaeffer, Alan C. Evans}

@@

\boutiques does not specify a particular language to compose
descriptors in a workflow, due to the large amount of specialized
frameworks to do so. However, workflows can be both composed from and
described as \boutiques descriptors: workflow engines can leverage the
\texttt{bosh} tools to call \boutiques tools from their descriptors;
in turn, workflows can be described as \boutiques descriptors provided
that the execution engine is containerized. Such a ``task
encapsulation'' model allows for a scalable and reliable execution of
workflows expressed in a variety of languages, as detailed in
~\cite{glatard2017software}.

@@

Applications are installed in a container image complying to the
Docker, Singularity or rootfs format. We intentionally support
multiple container formats as we anticipate that they will be used for
different purposes. For instance, Docker is well suited for developers
and users who manipulate applications on their local workstations or the
cloud. It is well documented, maintained and it has a rich ecosystem of
tools to help build and run containers on most operating systems. Singularity
is more suited for users and platforms that need to run applications
on shared computing clusters. Bridges exist among these containers
formats to convert container images across frameworks. For instance, a
platform dedicated to high-performance computing may accept descriptors
referring to Docker containers to facilitate application integration
by developers, and internally convert container images to Singularity
to run applications efficiently on clusters.

@@

\subsubsection{Flywheel}
%% gk: shall we check with them, and if nothing, remove?
\todo{Gunnar, Nathanael, John}

@@

The \boutiques descriptor specification allows describing a wide range
of applications, but it is also getting increasingly complex through
additions such as invocation schemas and dependencies among
inputs. Extending the \boutiques descriptor has two main goals. First,
accurately describing applications improves validation: incorrect input
values and execution results are more precisely detected when the
application descriptor is comprehensive, which is essential to improve
the reliability and user-friendliness of execution
platforms. Secondly, a rich descriptor schema reduces the need for
custom wrappers to integrate applications as even convoluted command
lines can be described: this is particularly interesting for
applications installed in containers since adding a wrapper to the
execution stack requires an update in the container image.

Nonetheless, a complex descriptor schema has a cost for application
developers and platforms, which we address as follows. For developers,
we maintain the set of mandatory descriptor properties as small as
possible so that simple applications can be described in a few lines
only (see Listing~\ref{listing:minimal}). For platforms, we aim at
supporting as many features as possible in the executor so that
only the following steps need to be implemented to support \boutiques
in a platform, regardless of the complexity of the descriptor:
\begin{enumerate}
  \item Input entry: generate the interface to allow users to enter
    input values.
  \item Input validation: create a JSON object from the interface,
    validate it against the invocation schema.
  \item Execution: generate and run the command line from the local
    executor and input JSON object.
  \item Output delivery: from the descriptor, identify the output files
    and deliver them to the user.
\end{enumerate}
In particular, advanced validation features such as dependencies
between inputs are embedded in the regular JSON validation performed
at step 2, without requiring the platform to support the related
descriptor properties.
\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=0pt,
               tabsize=4]{js}
{
  "name": "echo",
  "tool-version": "1.0",
  "description": "A simple script to test output files",
  "command-line": "echo [PARAM] > output.txt",
  "schema-version": "0.4",
  "inputs": [{
      "id": "param",
      "name": "Parameter",
      "value-key": "[INPUT_FILE]",
      "type": "File"
  }],
  "output-files": [{
      "id": "output_file",
      "name": "Output file",
      "path-template": "output.txt"
  }]

}
\end{minted}
\caption{A minimal \boutiques descriptor.} 
\label{listing:minimal}
\end{listing}

@@

The parallelization framework described in~\ref{sec:parallelization}
supports workflows by considering them as both applications and
platforms, which we think is a powerful but underexploited
model. Workflows are applications because they are invoked like any
other script or executable, possibly through a command line. Workflows
are also platforms because they may invoke other applications that
might themselves be described in \boutiques. This model is powerful
because it shields the \boutiques specification from specific workflow
constructs and allows a wide range of workflow engines to be described
uniformly.

\todo{CBRAIN's custom property might be standardized.}

@@

\boutiques helps reproduce\footnote{in Peng's
  sense~\cite{peng2011reproducible}} executions through containers and
formal command-line descriptions. With \boutiques, complete sets of
applications could be easily migrated across execution platforms,
including high-performance computing platforms and individual laptops,
to reproduce analyses. The \boutiques descriptor describes the
parameters and implementation of the application, and the invocation
schema describes the parameter values.

@@

In addition, important runtime parameters, for instance related to
multi-threading or available resources (storage, RAM), may be set by
the execution platform without being specified in the \boutiques
descriptor. Such runtime parameters may impact reproducibility in some
cases. To properly cover this issue, \boutiques descriptors should be
complemented by a provenance framework that captures a detailed trace
of the execution.

At a higher level, reproducibility is impacted by factors
such as data sharing~\cite{poldrack2014making} which are out of the
scope of \boutiques.

@@

So far, \boutiques has focused on the description of non-interactive
command-line applications. While such applications cover a large
fractions of the tools involved in scientific data processing, other
types of programs exist such as web services, interactive tools, and
applications with a graphical user interface (GUI).  Such application
classes could be described in \boutiques through a command-line
mapping. For instance, web services may be wrapped as command-line
applications using tools such as \texttt{curl} or
\texttt{wget}. Interactive applications may also be transformed to
non-interactive ones through configuration files. Finally, nothing
prevents a \boutiques tool from popping up a window for a user to
provide input through a GUI. This should, however, be specified
as an extension to the descriptor since most platforms would not support
this feature by default. Graphical output produced by applications
executed in containers should also be treated specifically.

@@

A few limitations remain that should be addressed in the
future. First, \boutiques moves the application porting bottleneck
from integration to validation. Using \boutiques, functions can be
automatically exported from frameworks such as Nipype and SPM,
creating hundreds of richly described applications potentially usable
by end-users. However, it remains unclear how to validate that these
applications actually execute as expected beyond executing packaged tests
distributed alongside a given tool. A \boutiques-specific testing
framework could be designed and potentially fed by existing frameworks,
to address this issue.

Another limitation is related to the security of containerized
applications. Since containers are usually controlled by application
developers rather than platforms, which is a good thing to reduce
application porting bottlenecks, nothing prevents developers to embed
malicious code in their container at any stage of the process,
possibly after a platform administrator inspected the
container. Containers are currently bulky file archives that are
cumbersome to inspect. Tools need to be developed to allow for an
easier characterization of container contents, for instance through
comparison digests with respect to validated base images. Singularity
containers have reduced security risks as compared to Docker containers,
but the issue of content transparency is still not avoided.

@@

\todo{BIDS apps}

The Common Workflow Language\footnote{\url{http://www.commonwl.org}}
is the piece of related work most closely related to
\boutiques. Several other application description frameworks have been
developed but they do not support containers.

@@

The Command Line Tool Description of the Common Workflow Language
(CWL) overlaps with the \boutiques descriptor. According to the history
of the CWL and \boutiques Github repositories, CWL started 6 months
before \boutiques (September 2014 vs. May 2015). The differences
highlighted below were identified from version 1.0 of the CWL Command
Line Tool
Description\footnote{\url{http://www.commonwl.org/v1.0/CommandLineTool.html}}.

@@

CWL has a formal command-line definition while \boutiques is more
flexible. CWL specifies command lines using an array containing an
executable and a set of arguments whereas \boutiques only uses a
string template. \boutiques' template approach may create issues in
some cases, but it also allows developers to add simple operations to
an application without having to write a specific wrapper. For
instance, a \boutiques command line may easily include input
decompression using, e.g., the tar command in addition to the main
tool command. On the contrary, CWL applications may only have a single
executable. Importantly, \boutiques' template system allows supporting
configuration files (see Section~\ref{sec:configuration-files}).

CWL uses ontologies while \boutiques does not. Ontologies allow for
richer definitions but they also have an overhead. The main
consequences are the following:
\begin{itemize}
\item CWL uses a specific framework for
validation, called SALAD (Semantic Annotations for Linked Avro Data)
whereas \boutiques uses plain JSON schema. The main goal of SALAD is to
allow ``working with complex data structures and document formats, such
as schemas, object references, and namespaces''. \boutiques only relies
on the basic types required to describe and validate a command line
syntactically. While the use of SALAD certainly allows for
higher-level validation, and may simplify the composition and
validation of complex workflows, it also introduces a substantial
overhead in the specification, and platforms have to use the validator
provided by CWL. On the contrary, a regular JSON validator can be used in \boutiques.
\item  CWL has a rich set of types whereas \boutiques only has simple
types. Again, this may be seen as a feature or as an overhead
depending on the context. \boutiques tries to limit the complexity of
the specification to facilitate its support by platforms where tools
will be integrated.
\end{itemize}

@@

\begin{itemize}
\item CWL tools have to write in a specific set of directories called
``designated output directory'', ``designated temporary directory''
and ``system temporary directory''. Tools are informed of the location
of such directories through environment variables. Having to write in
specific directories is problematic because tools have to be modified
to enable that. In \boutiques, the path of output files is defined
using a dedicated property.
\item CWL types are richer, not only
semantically but also syntactically. For instance, files have
properties for basename, dirname, location, path, checksum, etc.
\item \boutiques supports various types of containers (Docker, Singularity,
rootfs) while CWL supports only Docker. Both tools have rich requirements:
for instance, they may include RAM, disk usage, and walltime estimate; 
whereas \boutiques also supports number of requested CPU cores. CWL has hints, i.e.,
recommendations that only lead to warnings when not respected.
\item In \boutiques, dependencies can be defined among inputs, e.g., to
specify that an input may be used only when a particular flag is
activated. This is a very useful feature to improve validation, in
particular when tools have a lot of options.
\item In \boutiques, named
groups of inputs can be defined, which improves the presentation of
long parameter lists for the user and enables the definition of more
constraints within groups (e.g. mutually exclusive inputs).
\end{itemize}

\subsubsection{Other minor differences}

The \boutiques and CWL schemas have other minor differences that we
summarized in a public
spreadsheet\footnote{\url{https://docs.google.com/spreadsheets/d/1eLhpxzaaMPCIUFmV5Trs-WtaMpBO-c6QDQsmhwP7QIE/edit\#gid=0}}.

@@

Several frameworks have been created to facilitate the
integration of command-line applications in platforms. In
neuroinformatics, many platforms define a formal interface to embed
command-line applications. Among them, the Common
Toolkit\footnote{\url{http://www.commontk.org}} interoperates with
several platforms such as 3D Slicer~\cite{pieper20043d},
NiftyView~\cite{Craddock2016}, GIMIAS~\cite{larrabide2009gimias},
MedInria~\cite{larrabide2009gimias}, MeVisLab~\cite{heckel2009object}
and MITK workbench~\cite{nolden2013medical}. The framework, however,
remains tightly bound to the Common Toolkit's C++ implementation which
limits its adoption, e.g., in web platforms \todo{Fr\'ed\'eric, maybe
  you have more to say on this}.

In the distributed computing community, systems were also proposed to
facilitate the embedding of applications in platforms. The Grid
Execution Management for Legacy Code Architecture
(GEMLCA~\cite{delaitre2005gemlca}) was used to wrap applications in
grid computing systems. Interestingly, it has been used to embed
workflow engines in the SHIWA
platform~\cite{terstyanszky2014enabling}, in a similar but different
way than proposed by \boutiques (see Section~\ref{sec:parallelization}).

@@

\todo{FLI}

>>>>

\author[1]{Tristan Glatard}
\author[2,3]{Gregory Kiar}
\author[2,3]{Tristan Aumentado-Armstrong}
\author[2,3]{Natacha Beck}
\author[4]{Pierre Bellec}
\author[2,3]{R\'emi Bernard}
\author[5]{Sorina Camarasu-Pop}
\author[5]{Fr\'ed\'eric Cervenansky}
\author[2,3]{Samir Das}
\author[6]{Rafael Ferreira da Silva}
\author[7]{Guillaume Flandin}
\author[5]{Pascal Girard}
\author[8]{Krzysztof J. Gorgolewski}
\author[9]{Charles R.G. Guttmann}
\author[1]{Val\'erie Hayot-Sasson}
\author[4]{Pierre-Olivier Quirion}
\author[2,3]{Pierre Rioux}
\author[10]{Marc-\'Etienne Rousseau}
\author[2]{Alan C. Evans}

\affil[1]{Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada}
\affil[2]{McGill University, Montreal, Canada}
\affil[3]{Montreal Neurological Institute, Montreal, Canada}
\affil[4]{Centre de Recherche de l'Institut de G\'eriatrie de Montr\'eal CRIUGM, Montréal, QC, Canada}
\affil[5]{University of Lyon, CNRS, INSERM, CREATIS, Villeurbanne, France}
\affil[6]{University of Southern California, Information Sciences Institute, Marina del Rey, CA, USA}
\affil[7]{\todo{to be added}}
\affil[8]{Department of Psychology, Stanford University, Stanford, California, USA}
\affil[9]{Partners MS Center, Brigham and Women's Hospital, Brookline, Massachusetts, USA}
\affil[10]{Compute Canada}

@@

\boutiques does not specify a particular language to compose
descriptors in a workflow, due to the large amount of specialized
frameworks to do so. However, workflows can be both composed from and
described as \boutiques descriptors: workflow engines can leverage the
\texttt{bosh} tools to call \boutiques tools from their descriptors;
in turn, workflows can be described as \boutiques descriptors provided
that the execution engine is containerized. Such a ``task
encapsulation'' model allows for a scalable and reliable execution of
workflows expressed in a variety of languages, as detailed in
~\cite{GLATARD2017239}.

@@

Applications are installed in a container image complying to the
Docker, Singularity or rootfs format. We intentionally support
multiple container formats as we anticipate that they will be used for
different purposes. For instance, Docker is well suited for developers
and users who manipulate applications on their local workstations or
the cloud. It is well documented, maintained and it has a rich
ecosystem of tools to help build and run containers on most operating
systems. Singularity is more suited for users and platforms that need
to run applications on shared computing clusters. Bridges exist among
these containers formats to convert container images across
frameworks. For instance, a platform dedicated to high-performance computing may accept
descriptors referring to Docker containers to facilitate application
integration by developers, and run these images on clusters using
Singularity.

@@

@@

The \boutiques descriptor specification allows describing a wide range
of applications, but it is also getting increasingly complex through
additions such as invocation schemas and dependencies among
inputs. Extending the \boutiques descriptor has two main goals. First,
accurately describing applications improves validation: incorrect
input values and execution results are more precisely detected when
the application descriptor is comprehensive, which is essential to
improve the reliability and user-friendliness of execution
platforms. Secondly, a rich descriptor schema reduces the need for
custom application wrappers, for instance to create configuration
files from command-line options, as even convoluted command lines can
be described. This is particularly useful for containerized
applications since adding a wrapper requires an update of the
container image.

Nonetheless, a complex descriptor schema has a cost for application
developers and platforms, which we address as follows. For developers,
we maintain the set of mandatory descriptor properties as small as
possible so that simple applications can be described in a few lines
only (see Listing~\ref{listing:minimal}). For platforms, we aim at
supporting as many features as possible in the \texttt{bosh} executor
so that only the following steps need to be implemented to support
\boutiques in a platform, regardless of the complexity of the
descriptor:
\begin{enumerate}
  \item Input entry: generate the interface to enter inputs.
  \item Input validation: create a JSON invocation from the interface,
    validate it against the invocation schema.
  \item Input delivery: transfer the input files to the application
    execution location.
  \item Execution: pass the invocation to the \texttt{bosh} executor to run the application.
  \item Output delivery: from the descriptor, identify the output files
    and deliver them to the user.
\end{enumerate}
In particular, advanced validation features such as dependencies
between inputs are embedded in the regular JSON validation performed
at step 2, without requiring the platform to support the related
descriptor properties.
\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=0pt,
               tabsize=4]{js}
{
  "name": "echo",
  "tool-version": "1.0",
  "description": "A simple script to test output files",
  "command-line": "echo [PARAM] > output.txt",
  "schema-version": "0.5",
  "inputs": [{
      "id": "param",
      "name": "Parameter",
      "value-key": "[PARAM]",
      "type": "Number"
  }],
  "output-files": [{
      "id": "output_file",
      "name": "Output file",
      "path-template": "output.txt"
  }]
}
\end{minted}
\caption{A minimal \boutiques descriptor.} 
\label{listing:minimal}
\end{listing}

@@

\boutiques intentionally does not provide a workflow language to
compose tools, as this is already offered by several workflow
engines. Workflows can be either composed from \boutiques tools, as we
illustrated with the Nipype and MOTEUR engines, or described as
\boutiques applications, as we demonstrated with the Niak fMRI
pre-processing pipeline. Furthermore, the CBRAIN platform has a
sub-tasking mechanism that allows \boutiques tools to submit tasks,
which is used to parallelize workflows wrapped as \boutiques
tools. Based on the CBRAIN experience, we may specify the sub-tasking
mechanism in \boutiques so that other platforms can benefit from it.
This model is powerful because it shields the \boutiques specification
from specific workflow constructs and allows a wide range of workflow
engines to be described and used uniformly.

@@

\boutiques helps computational reproducibility, in Peng's
sense~\cite{peng2011reproducible}, through containers and formal
command-line descriptions. With \boutiques, complete sets of
applications could be easily migrated across execution platforms,
including high-performance computing platforms and individual laptops,
to reproduce analyses. The \boutiques descriptor describes the
parameters and implementation of the application, and the invocation
schema describes the parameter values.

@@

In addition, important runtime parameters, for instance related to
multi-threading or available resources (storage, RAM), may be set by
the execution platform without being specified in the \boutiques
descriptor. Such runtime parameters may impact reproducibility in some
cases. To properly cover this issue, \boutiques descriptors should be
complemented by a provenance framework that captures a detailed trace
of the execution. We plan to leverage the provenance format being
defined by the Neuro-Imaging Data Model
Workflow~\cite{ghosh2017capturing} initiative for this purpose.

@@

So far, \boutiques has focused on the description of non-interactive
command-line applications. While such applications cover a large
fraction of the tools involved in scientific data processing, other
types of programs exist such as web services, interactive tools, and
applications with a graphical user interface (GUI).  Such application
classes could be described in \boutiques through a command-line
mapping. For instance, web services may be wrapped as command-line
applications using tools such as \texttt{curl} or
\texttt{wget}. Interactive applications may also be transformed to
non-interactive ones through configuration files. Finally, nothing
prevents a \boutiques tool from popping up a window for a user to
provide input through a GUI. This should, however, be specified
as an extension to the descriptor since most platforms would not support
this feature by default. Graphical output produced by applications
executed in containers should also be treated specifically.

@@

A few limitations remain that should be addressed in the
future. First, \boutiques moves the application porting bottleneck
from integration to validation. Using \boutiques, functions can be
automatically exported from frameworks such as Nipype and SPM,
creating hundreds of richly described applications potentially usable
by end-users. However, the automated validation of such applications
remains challenging. A \boutiques-specific testing framework could be
designed and potentially fed by existing frameworks, to address this
issue.

Another limitation is related to the security of containerized
applications. Since containers are usually controlled by application
developers rather than platforms, which is a good thing to reduce
application porting bottlenecks, nothing prevents developers to embed
malicious code in their container at any stage of the process,
possibly after a platform administrator inspected the
container. Containers are bulky file archives that are
cumbersome to inspect. Tools need to be developed to allow for an
easier characterization of container contents, for instance through
comparison digests with respect to validated base images. Singularity
containers have reduced security risks as compared to Docker containers,
but the issue of content transparency is still not avoided.

@@

Several frameworks have been developed to describe and integrate
applications in various types of platforms. \boutiques focuses on (1)
fully-automatic integration of applications, including deployment on
heterogeneous computing resources through containers, (2)
comprehensive input validation through a strict JSON schema, (3)
flexible application description through a rich JSON
schema.

@@

The Common Workflow Language
(CWL\footnote{\url{http://www.commonwl.org}}) is the piece of related
work most closely related to \boutiques as it provides a formal way to
describe containerized applications. In particular, CWL's Command Line
Tool Description overlaps with the \boutiques descriptor. This Section
highlights the main differences between CWL and \boutiques, based on
version 1.0 of the CWL Command Line Tool
Description\footnote{\url{http://www.commonwl.org/v1.0/CommandLineTool.html}}. According
to GitHub, CWL started 6 months before \boutiques (September 2014
vs. May 2015).

@@

CWL imposes a strict command-line format while \boutiques is more
flexible. CWL specifies command lines using an array containing an
executable and a set of arguments whereas \boutiques only uses a
string template. \boutiques' template approach may create issues in
some cases, but it also allows developers to add simple operations to
an application without having to write a specific wrapper. For
instance, a \boutiques command line may easily include input
decompression using the \texttt{tar} command in addition to the main
tool command. Importantly, \boutiques' template system allows
supporting configuration files.

CWL uses ontologies while \boutiques does not. Ontologies allow for
richer definitions but they also have an overhead. The main
consequences are the following:
\begin{itemize}
\item CWL uses a specific framework for
validation, called SALAD (Semantic Annotations for Linked Avro Data)
whereas \boutiques uses plain JSON schema. The main goal of SALAD is to
allow ``working with complex data structures and document formats, such
as schemas, object references, and namespaces''. \boutiques only relies
on the basic types required to describe and validate a command line
syntactically. While the use of SALAD certainly allows for
higher-level validation, and may simplify the composition and
validation of complex workflows, it also introduces a substantial
overhead in the specification, and platforms have to use the validator
provided by CWL. On the contrary, a regular JSON validator can be used in \boutiques.
\item  CWL has a rich set of types whereas \boutiques only has simple
types. This may again be seen as a feature or as an overhead
depending on the context. \boutiques tries to limit the complexity of
the specification to facilitate its support by platforms where tools
will be integrated.
\end{itemize}

@@

\begin{itemize}
\item CWL tools have to write in a specific set of directories called
``designated output directory'', ``designated temporary directory''
and ``system temporary directory''. Tools are informed of the location
of such directories through environment variables. Having to write in
specific directories is problematic because tools have to be modified
to enable that. In \boutiques, the path of output files is defined
using a dedicated property.
\item CWL types are richer, not only
semantically but also syntactically. For instance, files have
properties for basename, dirname, location, path, checksum, etc.
\item \boutiques supports various types of containers (Docker,
  Singularity, rootfs) while CWL supports only Docker. Both tools have
  rich requirements: for instance, they may include RAM, disk usage,
  and walltime estimate. CWL has hints, i.e., recommendations that
  only lead to warnings when not respected.
\item In \boutiques, dependencies can be defined among inputs, e.g., to
specify that an input may be used only when a particular flag is
activated. This is a very useful feature to improve validation, in
particular when tools have a lot of options.
\item In \boutiques, named
groups of inputs can be defined, which improves the presentation of
long parameter lists for the user and enables the definition of more
constraints within groups (e.g. mutually exclusive inputs).
\end{itemize}

\subsection{BIDS apps}

BIDS apps~\cite{gorgolewski2017bids} specify a framework for
neuroimaging applications to process datasets complying to the Brain
Imaging Data Structure (BIDS). They share common goals with
\boutiques, in particular reusability across platforms through
containerization. Conceptually, however, BIDS apps and \boutiques are
different since BIDS apps intend to standardize application interfaces
while \boutiques intends to describe them as flexibly as
possible. BIDS apps have a specific set of inputs and outputs, for
instance the input dataset, that have to be present in a specific
order on the command line for the application to be valid. The
specification adopted by BIDS apps simplifies the integration of
applications in platforms as they all comply to the same
interface. However, it is also limited to the subset of neuroimaging
applications that process BIDS datasets and it does not formally
describe application-specific inputs. All in all, BIDS apps and
\boutiques complement each other: BIDS apps provide a practical way to
integrate neuroimaging applications, while \boutiques offers a formal
description of their specific parameters.  \boutiques descriptors can
be generated from BIDS apps using the \texttt{bosh} importer.

@@

Several other frameworks have been created to facilitate the
integration of command-line applications in platforms. In
neuroinformatics, many platforms define a formal interface to embed
command-line applications. Among them, the Common
Toolkit\footnote{\url{http://www.commontk.org}} interoperates with
several platforms such as 3D Slicer~\cite{pieper20043d},
NiftyView~\cite{Craddock2016}, GIMIAS~\cite{larrabide2009gimias},
MedInria~\cite{larrabide2009gimias}, MeVisLab~\cite{heckel2009object}
and MITK workbench~\cite{nolden2013medical}. The framework, however,
remains tightly bound to the Common Toolkit's C++ implementation which
limits its adoption, e.g., in web platforms.

In the distributed computing community, systems were also proposed to
facilitate the embedding of applications in platforms. The Grid
Execution Management for Legacy Code Architecture
(GEMLCA~\cite{delaitre2005gemlca}) was used to wrap applications in
grid computing systems. Interestingly, it has been used to embed
workflow engines in the SHIWA
platform~\cite{terstyanszky2014enabling}, in a similar but different
way than proposed by \boutiques.

@@

Pipeline integration for the MICCAI 2016 challenge was funded by the
French National Agency for Research (ANR) through ``France
Life-Imaging''. We also thank Compute Canada and Calcul Québec for
providing a computing infrastructure supporting Docker and Singularity
containers. This research was undertaken thanks in part to funding
from the Canada First Research Excellence Fund, awarded to McGill
University for the Healthy Brains for Healthy Lives initiative.
